---
title: "Heart Disease Analysis – Typical Angina"
author: "Massimiliano Di Giuseppe"
date: "2026-04-25"
output: html_document
---

# Introduction

This analysis investigates which variables distinguish healthy and diseased individuals among patients presenting with typical angina.

The goal is to identify the most informative predictors while controlling for symptom type.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)

df <- read.csv("https://raw.githubusercontent.com/mrdbourke/zero-to-mastery-ml/master/data/heart-disease.csv")
```



```{r data-cleaning}
df_clean <- df %>%
  mutate(
    sex = ifelse(sex == 1, "Maschio", "Femmina"),
    target = ifelse(target == 1, "Malattia", "Sano"),
    cp = case_when(
      cp == 0 ~ "Angina tipica",
      cp == 1 ~ "Angina atipica",
      cp == 2 ~ "Dolore non anginoso",
      cp == 3 ~ "Asintomatico",
      TRUE ~ NA_character_
    ),
    fbs = ifelse(fbs == 1, ">120 mg/dl", "≤120 mg/dl"),
    exang = ifelse(exang == 1, "Sì", "No")
  ) %>%
  rename(
    eta = age,
    sesso = sex,
    pressione_riposo = trestbps,
    colesterolo = chol,
    glicemia_digiuno = fbs,
    ecg_riposo = restecg,
    freq_cardiaca_max = thalach,
    angina_sforzo = exang,
    depressione_st = oldpeak,
    pendenza_st = slope,
    vasi_principali = ca,
    talassemia = thal,
    tipo_dolore = cp,
    diagnosi = target
  ) %>%
  mutate(
    malattia_binaria = ifelse(diagnosi == "Malattia", 1, 0)
  )
```

```{r filtering}
df_angina_tipica <- df_clean %>%
  filter(tipo_dolore == "Angina tipica")
```

# Logistic Regression Model

```{r logistic-model}
model <- glm(
  malattia_binaria ~ 
    pressione_riposo +
    colesterolo +
    freq_cardiaca_max +
    depressione_st +
    vasi_principali +
    pendenza_st +
    ecg_riposo +
    sesso,
  data = df_angina_tipica,
  family = binomial
)

summary(model)
exp(coef(model))
```

## Results

The analysis shows that vascular involvement (number of major vessels) and sex are the strongest predictors of disease.

ST slope shows a moderate association, while traditional cardiovascular variables such as cholesterol and resting blood pressure do not significantly discriminate between healthy and diseased individuals.

This suggests that the relationship between variables is not purely linear and depends on the clinical context.

## Conclusion

This analysis highlights the importance of multivariate approaches in medical data.

Single variables alone are often insufficient, and meaningful insights emerge only when multiple factors are considered together.